Effects of rotation and magnetic braking on Li abundances for solar-type stars

### 2019-07-11 project/overleaf/paper.tex
Paper v3.1

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\raggedbottom
\setlength{\parskip}{1em}
%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of the influence of rotation and magnetic braking (MB) on lithium (Li) depletion during pre-main sequence (PMS) and main sequence (MS) for solar-type stars.
\newline\textit{Methods.} The effects of rotational mixing and of the hydrostatic effects of rotation on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB one (with magnetic field intensities ranging between 3.5 and 5.5 G) and the data obtained from the two set of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS must if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $T_{Li} \approx 2.5 x 10^6\; K$, and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. Solar-type stars are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS), when the convection zone retreats and the temperature at BCZ is cooler than $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances, in addition to showing a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit to the star surface. However different abundances of Li have been observed for different star populations (\citet{Somers2014} and references therein).\par

From approximately the 1920s of the last century it has been accepted that rotation is capable of inducing the mixing of the elements present inside the stars. The studies of Eddington (1925) and Vogt (1925) had revealed the fact that the southern circulation must necessarily arise in those stars in rotation as a consequence of the von Zeipel paradox: a star cannot be simultaneously in thermal and hydrostatic equilibrium if its rotation speed depends exclusively on its radius. The solution to this depends on the existence of meridian circulations. At the same time, this type of instability implies an impact on the convection mechanisms not considered by the standard models. Today it is known that in combination with this type of instability exists a catalogue of others that appear as a consequence of assuming models in rotation \citep{Maeder2003a}, especially the differential rotation as a function of the radius of the star or the deviation of the spherical symmetry. Although initially these effects could be candidates to explain the observations, the fact is that the Sun is a slow rotator, meaning slow rotation that one for which the maximum centrifugal force is much smaller than the force of gravity, its deviation of the measured spherical symmetry is of the order of $10^{-5}$ and the time scale for the southern circulation is of the order of $10^{12}$ years, a time period well above the nuclear combustion time scale \citep{Pinsonneault1997}. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

The impact of rotation on PMS and on Li depletion for solar-type stars has already been discussed (\citet{Pinsonneault1997}, \citet{Jeffries2004}, \citet{Somers2014}). These previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. As previously mentioned, it's thought that AML has a direct influence in the mixing processes and transport of the momentum could be produced by a number of mechanisms that could be important: mass loss, magnetic fields and waves (g-modes). Concerning the two latter, waves and magnetic fields have the property that they generally transmit angular momentum (AM) much more effectively than they induce mixing \citep{Denissenkov2007}. As result, the efficiency of these processes is most easily tested by their impact on AM evolution. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on it by MB as a consequence of the solidary rotation that produces on the particles that constitute the solar wind and the torque derived from this coupling. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. A proper consideration of the interaction between rotation and magnetic fields is therefore essential. We have adapted the one-dimensional stellar evolution code used in this work to model MB and include its effect on the equations for the evolution of the AM distribution and stellar structure. This produces a more complete, though still reasonably simple, model for AM evolution. In this study we are now focused in investigating not only the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars.\par

In the remainder of this paper, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar wind. In its simplest form, implementation of the model requires specification of both the surface magnetic field strength $B$ and the initial  rotation rate $\Omega$ at star surface. Our simulations trace the rotational history and A(Li) of a $1 M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par

We define here also some conventions and assumptions adopted in the paper. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, proto} = 0.0142$, not the current photospheric metallicity $Z = Z_{\odot, photosphere} = 0.0134$, as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $R_{\odot} = 6.957x10^{10} cm$ and $M_{\odot} = 1.988x10^{33} g$ which are consistent with IAU resolutions \citep{Mamajek2015} . For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. This work has been used as a starting point for our work regarding the parameterization of the MESA project calibrated to reproduce the measured values on the solar surface, in particular the section 4.1 of the aforementioned paper.\par

This paper is structured as follows: Section 2 gives an overview of rotation and magnetic fields physical processes and how do they impact on the star evolution. We also summarized the most relevant parameters configured in MESA for the model simulations. In Section 3, the model outputs are presented and discussed in detail. Finally, Section 4 finishes this paper with a discussion of open points and future works.\par

@@

@@

Angular velocity, or rather the way in which it evolves, also conditions the mechanism by which theoretical models solve the equations of stellar structure. It is evident that stars have a three-dimensional structure and for this reason the equations that model their evolution must take this fact into account. However, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two, if it is fulfilled that the angular velocity is constant in regions subjected to the same pressure (isobaric). This is what is known as shellular approximation \citep{Meynet1997} and is based on calculating the impact of the centrifugal acceleration caused by the rotation in the stellar evolution equations \citep{Endal1976}. MESA uses this simplification to accelerate computational calculations.\par

\subsection{Effects of rotation}
Rotation may affect the equations of stellar structure in four ways according to \citet{Endal1976}:
%\renewcommand{\theenumi}{\arabic{enumi}.}
\begin{enumerate}
    \item The centrifugal force causes the local effective gravity ($g_{eff}$) associated with a point of the star not located on the axis of rotation to be reduced.
    \item In general, the centrifugal force will not be parallel to the gravitational force and, as a consequence of this, the equipotential layers will cease to be spherical.
    \item As the von Zeipel theorem states, the radiative flux in a uniformly rotating star is proportional to the $g_{eff}$, therefore the radiative flux is not constant on an equipotential surface.
    \item Rotation can suppress certain mechanisms that trigger convection, indirectly affecting mixing processes.
\end{enumerate}

Observational data support the fact that stars are born with different angular velocities even if they come from the same interstellar cloud and that throughout evolution they undergo a deceleration process \citep{Skumanich1972} in the time period between 0.5 and 5 Gyr \citep{Somers2014a}. This range of different velocities implies that the mixing mechanisms induced by rotation vary their efficiency and effectiveness and therefore have a direct influence on the abundances of the different chemical elements observed on the surface of stars.  This fact, together with the differential degree of rotation to which the limits between the radiative and convective zones of the star are subjected, is of particular interest for our study since Li, as mentioned in the introduction, is destroyed at a low temperature, so variations in the intensity of the mixing processes can cause convective zones to extend to interior zones with temperatures higher than $T_{Li}$. The deceleration process to which a star is subjected throughout its evolution and the direct link that the rotation has with the mixing processes also offer a natural and consistent explanation for the fact that the destruction of Li is more accentuated or efficient in young stars than in old stars \citep{Sestito2005}. In a similar way, stellar rotation rates show a large dispersion at ZAMS \citep{Stauffer1984}, providing in this way a consistent argument for explaining the variance of Li abundances among stars which have a similar mass and were formed within the same cloud. Although this approach is not a novelty at all, it is no less true that in those studies based on it the effect caused by MB over the rotation on the PMS and MS is often overlooked. \par

In reference to MESA and as stated in \citet{Paxton2013}: "the transport of AM and chemical elements due to rotationally induced instabilities is implemented in a diffusion approximation \citet{Endal1978}. MESA calculates diffusion coefficients for five different rotationally induced mixing processes: dynamical shear instability, Solberg-H{\o}iland instability, secular shear instability, Eddington-Sweet circulation, and the Goldreich-Schubert-Fricke instability".

\subsection{Modelling magnetic braking}
The origin of magnetic fields is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields (see \citet{Dudorov2014} for details about the Theory of fossil magnetic field), or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, the presence of magnetic fields with intensities of the order of kG \citep{Hussain2014} has been a sine qua non condition for dealing with certain data observed in T Tauri type stars in young, accreting PMS systems \citep{Johns-Krull2007}. One of this features is the long rotating periods  which is much longer than predicted from AM conservation. A key question is to understand how these young stars can accrete large amounts of material from their circumstellar disk with high specific angular momentum and at the same time they're able to keep slow rotation periods, between 7-10 days \citep{Hussain2014}. One plausible explanation for this behaviour considers the AML produced by the interaction between the magnetic field and the particles that form its circumstellar disk \citep{Zanni2012}. Once the accretion process stops or becomes less efficient then the star is able to spin up as it contracts towards the ZAMS.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. In these cases, the dominant type of energy transport in the outer zones of the star is radiative. On the other hand, the less massive stars, such as those of the solar-type, have convective covers and radiative core. In both types of stars strong magnetic fields can be generated in the limits of both zones (the so-called tachocline) but with notable differences. In the case of massive stars, these magnetic fields can survive for a long time but not reach the surface of the star. On the other hand in less massive stars these magnetic fields, although they are able to reach the surface of the star, decay quickly in time, in a few hundred years or even only in a few decades \citep{Chabrier2006}, so it is necessary a mechanism for generating magnetic field, e.g. a dynamo type, continued in time to explain its existence. The way in which this dynamo process is able to generate magnetic fields depends on different parameters and to this day this mechanism is still not fully accepted and is subject of controversy, e.g. \citep{Charbonneau2010}.\par

In the present work we do not take into account the existence of possible magnetic fields during the T-Tauri phase and we adopt a simple theoretical approach to establish their existence during the ZAMS approximation phase. The adopted criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine will be activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participate in the star AML. Additionally, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches TAMS.\par

Our main objective for this paper is to check, in a first approximation, the degree of validity of an approach that includes MB as a complementary mechanism of AML that can contribute to explain the evolution of Li in solar-type stars. More advanced models of the arrangement of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\; G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology. This change of typology in the magnetic field, from poloidal to toroidal, is known as the $\omega$-effect. On the other hand, we have that the opposite case can also occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible, on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and, on the other, to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it. Under these circumstances, the magnetic field lines will be pushed to higher areas of the CZ as the material moves vertically and, due to the combined effect of stellar rotation plus Coriolis forces, the magnetic lines will be twisted and give rise to a poloidal magnetic field. This is what is known as the $\alpha$-effect.\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\; G$, twice the value of the Earth's one. Additionally its topology is varied and complex. As mentioned in the previous section, the dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on they reach the solar surface. Among them, we can find sunspots dominated by magnetic fields with an approximate intensity of $B\approx10^3\; G$ and coronal loops with field strengths of $B\approx10^2\; G$ at the photosphere, and $B\approx10 G$ in larger coronal heights \citep{Aschwanden2014}. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism (see, e.g. \citet{Brandenburg2004}, \citet{Charbonneau2010} or \citet{Brun2017}). Whatever the origins of surface magnetic fields are, it's expected to couple to the wind mass-loss and, if this is strong enough, to produce MB \citet{UdDoula2002}, \citet{Ud-Doula2007}, \citet{Ud-Doula2008} \citet{Meynet2010}.\par

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{nccmath}
\usepackage{isotope}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[firstpage]{draftwatermark}
\SetWatermarkText{DRAFT}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\raggedbottom
\setlength{\parskip}{1em}
%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the hydrostatic effects of rotation on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB one (with magnetic field intensities ranging between 3.5 and 5.5 G) and the data obtained from the two set of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS must if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $T_{Li} \approx 2.5 x 10^6\; K$, and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances, in addition to showing a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit to the star surface. However, different abundances of Li have been observed for different star populations (\citet{Somers2014} and references therein).\par

From approximately the 1920s of the last century it has been accepted that rotation is capable of inducing the mixing of the elements present inside the stars. The studies of Eddington (1925) and Vogt (1925) had revealed the fact that the southern circulation must necessarily arise in those stars in rotation as a consequence of the von Zeipel paradox: a star cannot be simultaneously in thermal and hydrostatic equilibrium if its rotation speed depends exclusively on its radius. The solution to this depends on the existence of southern circulations. At the same time, this type of instability implies an impact on the convection mechanisms not considered by the standard models. Today it is known that in combination with this type of instability exists a catalogue of others that appear as a consequence of assuming models in rotation \citep{Maeder2003a}, especially the differential rotation as a function of the radius of the star or the deviation of the spherical symmetry. Although initially these effects could be candidates to explain the observations, the fact is that the Sun is a slow rotator, meaning slow rotation that one for which the maximum centrifugal force is much smaller than the force of gravity. Furthermore, its deviation of the measured spherical symmetry is of the order of $10^{-5}$ and the time scale for the southern circulation is of the order of $10^{12}$ years, a time period well above the nuclear combustion time scale \citep{Pinsonneault1997}. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

The impact of rotation on PMS and on Li depletion for solar-type stars has already extensively debated in the past (\citet{Pinsonneault1997}, \citet{Jeffries2004}, \citet{Somers2014}). However, these previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It's thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and waves (g-modes). Concerning the two latter, waves and magnetic fields have the property that they generally transmit angular momentum (AM) much more effectively than they induce mixing \citep{Denissenkov2007}. As result, the efficiency of these processes is most easily tested by their impact on AM evolution. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on it by MB as a consequence of the solidary rotation that produces on the particles that constitute the solar wind and the torque derived from this coupling. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. A proper consideration of the interaction between rotation and magnetic fields is therefore essential. We have adapted the one-dimensional stellar evolution code used in this work to model MB and included its effect on the equations for the evolution of the AM distribution and stellar structure. This produces a more complete, though still reasonably simple, model for AM evolution. It's worth to mention that in this study we don't only focused in investigating the effects of rotation on the Li content during PMS and MS but also on the indirect role played by the MB on the Li destruction as a result of its influence on the rotational history of solar-type stars.\par

\subsection{Effects of rotation}
Rotation may affect the equations of stellar structure in four different ways according to \citet{Endal1976}:
%\renewcommand{\theenumi}{\arabic{enumi}.}
\begin{enumerate}
    \item The centrifugal force causes the local effective gravity ($g_{eff}$) associated with a point of the star not located on the axis of rotation to be reduced.
    \item In general, the centrifugal force will not be parallel to the gravitational force and, as a consequence of this, the equipotential layers will cease to be spherical.
    \item As the von Zeipel theorem states, the radiative flux in a uniformly rotating star is proportional to the $g_{eff}$, therefore the radiative flux is not constant on an equipotential surface.
    \item Rotation can suppress certain mechanisms that trigger convection, indirectly affecting mixing processes.
\end{enumerate}

Observational data support the fact that stars are born with different angular velocities even if they come from the same interstellar cloud and that throughout evolution they undergo a deceleration process \citep{Skumanich1972} in a time period between 0.5 and 5 Gyr \citep{Somers2014a}. This range of different velocities implies that the mixing mechanisms induced by rotation vary their efficiency and effectiveness and therefore have a direct influence on the abundances of the different chemical elements observed on the surface of stars.  This fact, together with the differential degree of rotation to which the limits between the radiative and convective zones of the star are subjected, is of particular interest for our study since Li, as mentioned in the introduction, is destroyed at a low temperature. Therefore, variations in the intensity of the mixing processes can cause convective zones to extend to interior zones with temperatures higher than $T_{Li}$. The deceleration process to which a star is subjected throughout its evolution and the direct link that the rotation has with the mixing processes also offer a natural and consistent explanation for the fact that the destruction of Li is more accentuated or efficient in young stars than in old stars \citep{Sestito2005}. In a similar way, stellar rotation rates show a large dispersion at ZAMS \citep{Stauffer1984}, providing in this way a consistent argument for explaining the variance of Li abundances among stars which have a similar mass and were formed within the same cloud. Although this approach is not a novelty at all, it is no less true that in those studies based on it the effect caused by MB over the rotation on the PMS and MS is often overlooked. \par

In the remainder of this paper, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar wind. In its simplest form, the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial  rotation rate ($\Omega$) at star surface. Our simulations trace the rotational history and A(Li) of a $1\, M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par

We define here also some conventions and assumptions adopted in the paper. For the models presented in this paper we adopt solar-scaled abundances and assumes the \citet{Asplund2009} protosolar birth cloud metallicity ($Z = Z_{\odot, proto} = 0.0142$) not the current photospheric metallicity ($Z = Z_{\odot, photosphere} = 0.0134$) as the reference value. We also adopt the following nominal values to express stellar properties values in SI units $R_{\odot} = 6.957x10^{10}\, cm$ and $M_{\odot} = 1.988x10^{33}\, g$ which are consistent with IAU resolutions \citep{Mamajek2015}. For a detailed description of the physics adopted in this paper, refer to \citet{Choi2016}. That work, in particular its section 4.1, has been used as a starting point for ours regarding the parameterization of the MESA project which is calibrated to reproduce the measured values on the solar surface.\par

\subsection{Effects of magnetic fields}
The origin of magnetic fields is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields (see \citet{Dudorov2014} for details about the theory of fossil magnetic field), or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, the presence of magnetic fields with intensities of the order of kG \citep{Hussain2014} has been a sine qua non condition for dealing with certain data observed in T Tauri type stars in young, accreting PMS systems \citep{Johns-Krull2007}. One of this features is the long rotating periods  which is much longer than predicted from AM conservation. A key question is to understand how these young stars can accrete large amounts of material from their circumstellar disk with high specific angular momentum and at the same time they're able to keep slow rotation periods, between 7-10 days \citep{Hussain2014}. One plausible explanation for this behaviour considers the AML produced by the interaction between the magnetic field and the particles that form its circumstellar disk \citep{Zanni2012}. Once the accretion process stops or becomes less efficient then the star is able to spin up as it contracts towards the ZAMS.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. In these cases, the dominant type of energy transport in the outer zones of the star is radiative. On the other hand, the less massive stars, such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences. In the case of massive stars, these magnetic fields can survive for a long time but not reach the surface of the star. On the other hand in less massive stars these magnetic fields, although they are able to reach the surface of the star, decay quickly in time, in a few hundred years or even only in a few decades \citep{Chabrier2006}, so it is necessary a mechanism for generating magnetic field, e.g. a dynamo type, continued in time to explain its existence. The way in which this dynamo process is able to generate magnetic fields depends on different parameters and to this day this mechanism is still not fully accepted and is subject of controversy, e.g. \citep{Charbonneau2010}.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\, G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology. This change of typology in the magnetic field, from poloidal to toroidal, is known as the $\omega$-effect. On the other hand, we have that the opposite case can also occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible, on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and, on the other, to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it. Under these circumstances the magnetic field lines will be pushed to higher areas of the CZ as the material moves vertically and, due to the combined effect of stellar rotation plus Coriolis forces, the magnetic lines will be twisted and give rise to a poloidal magnetic field. This is what is known as the $\alpha$-effect.\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\, G$, twice the value of the Earth's one. Additionally its topology is varied and complex. As mentioned in the previous section, the dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on they reach the solar surface. Among them, we can find sunspots dominated by magnetic fields with an approximate intensity of $B\approx10^3\, G$ and coronal loops with field strengths of $B\approx10^2\, G$ at the photosphere, and $B\approx 10\, G$ in larger coronal heights \citep{Aschwanden2014}. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism (see, e.g. \citet{Brandenburg2004}, \citet{Charbonneau2010} or \citet{Brun2017}). Whatever the origins of surface magnetic fields are, it's expected to couple to the wind mass-loss and, if this is strong enough, to produce MB \citet{UdDoula2002}, \citet{Ud-Doula2007}, \citet{Ud-Doula2008} \citet{Meynet2010}.\par

\subsection{Effects of magnetic braking}
In its most widely accepted approach, the MB is linked to Skumanich's work in which he develops an empirical law of evolution of the MB according to the age of the star. To calibrate this law, Skumanich relied on data obtained from type G stars found in the MS \citep{Skumanich1972}. The influence of the MB on the evolution of the star is closely related to the transport of the AM and by extension, influenced by how the chemical elements are transported within it. According to this approach \citep{Meynet2010}, two main mechanisms are distinguished attending to how the transport of the AM is produced which influence the mixing of these elements:

Returning to section \ref{surf_mf} on the estimated mass loss ratio for a solar-type star, approximately $10^{-13}M_{\odot} \; yr^{-1}$, and using this value to estimate $\Dot{J}$, the result obtained can be considered relatively modest as to decisively influence the evolution of the star. On the other hand, if we consider that the convective movements that take place in the stellar interior together with the presence of differential rotation can give rise to the generation of magnetic fields that end up reaching the surface of the star \citep{Langer2012} and, the AML can be increased in several orders of magnitude as a consequence of the magnetic field that forces the ionized particles that compose the stellar wind to turn with the same angular speed and whose influence on the stellar wind extends a distance several times the radius of the star (see \citet{UdDoula2002}, \citet{Ud-Doula2007} and \citet{Ud-Doula2008} for more details).

This paper is structured as follows: Section 2 gives an overview of rotation and magnetic fields physical processes and how do they impact on the star evolution. We also summarized here the most relevant parameters configured in MESA for the model simulations. In Section 3, the model outputs are presented and discussed in detail. Finally, Section 4 finishes this paper with a discussion of open points and future works.\par

@@

Keeping in mind this approach, our major objective is aimed to check the degree of validity of the resulting model after including the MB effect as a complementary mechanism of AML that can contribute to explain the evolution of Li in the Sun and other solar-type stars. A set of different simulation scenarios with magnetic field strengths ranging from between 3 and 5.5 G have been exercised. We'd like to stress again that the magnetic field remains constant during the PMS \& MS phases. We compute the evolution of $1M_{\sun}$ stellar models at solar metallicity with $\Omega / \Omega_{crit}$ varying between $0.0084$ and $0.0336$. More advanced models for more complex arrangements of the magnetic field as well as the evolution of its intensity throughout the life of the star are left for later work.\par

The rest of this section is devoted to describe the most relevant aspects and assumptions made in modelling the evolution of rotation, magnetic braking and angular momentum in MESA.\par

@@

Angular velocity, or rather the way in which it evolves, also conditions the mechanism by which theoretical models solve the equations of stellar structure. It is evident that stars have a three-dimensional structure and for this reason the equations that model their evolution must take this fact into account. However, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two, if it is fulfilled that the angular velocity is constant in regions subjected to the same pressure (isobaric). This is what is known as shellular approximation \citep{Meynet1997} and is based on calculating the impact of the centrifugal acceleration caused by the rotation in the stellar evolution equations \citep{Endal1976}. MESA uses this simplification to accelerate computational calculations. Additionally, as stated in \citet{Paxton2013}: "the transport of AM and chemical elements due to rotationally induced instabilities is implemented in a diffusion approximation \citet{Endal1978}. MESA calculates diffusion coefficients for five different rotationally induced mixing processes: dynamical shear instability, Solberg-H{\o}iland instability, secular shear instability, Eddington-Sweet circulation, and the Goldreich-Schubert-Fricke instability".\par

\subsection{Modelling magnetic braking}
In the presence of mass-loss, AM is removed by the material in the stellar wind, resulting in a spin-down effect. This is quite important for example in massive stars, that are known to be rapidly rotating and have strong stellar winds. It has been also observed that strongly magnetic intermediate-mass stars typically have rotation rates much slower than other stars in their parent population \citep{Mathys2006}. In those stars, the presence of such magnetic field will interact with the mass loss and the Alfv\'{e}n radius ($R_{A}$) plays an important role in this interaction. The $R_{A}$ is defined as the point in which the magnetic field energy density $B^{2}/8\upi$ and the kinetic energy density $\frac{1}{2}\rho\nu^{2}$ of the stellar wind are balanced, that is, where the coefficient between both is equal to 1. In case that $R_{A}$ is greater than the star radius, then the wind flow will have to follow the magnetic field. As a consequence, the material will leave the stellar surface with a higher specific AM, as the co-rotation radius has increased and it will roughly correspond to the $R_{A}$. This effect is known as MB.\par

In the present work we do not take into account either the influence of internal magnetic fields or their existence during the T-Tauri phase. We adopt a simple and pragmatic theoretical approach to establish their occurrence when the star is reaching the ZAMS phase. The adopted criterion is based on the simultaneous existence of an extensive convective layer and a radiative core. From this moment on, the MB routine will be activated, acting as an additional mechanism to those existing in the MESA evolutionary code that participates in the star AML. Additionally, we assume that the magnetic field does not vary its intensity throughout the evolution of the star until it reaches the TAMS.\par

The energy ratio (eq. \ref{eq:wind_conf}) is denominated wind confinement magnetic parameter ($\eta_*$) and defines a characteristic parameter for the relative effectiveness of the magnetic fields in circumscribing and/or channeling the wind outflow \citep{UdDoula2002}. In turn, of the observable characteristics of the solar wind, we especially care about the parameters related to the amount of mass that is able to pull out of the outer layers of the star in a given time interval, and the speed that reaches the wind itself measured at a great distance from the star, the terminal velocity ($\nu_{esc}$).\par
